---
title: "p8105_hw3_jl6325"
author: "Jianing Liu"
date: "2023-10-10"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



# Problem 1

```{r}
data("instacart")

instacart = 
instacart |>
  as_tibble()
```


```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```


```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```


```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |>
  knitr::kable(digits = 2)
```


# Problem 2


```{r}
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 |>
  as.tibble()
```


```{r}
bs =
  brfss_smart2010 |>
  rename(state = Locationabbr,
         county = Locationdesc) |>
  janitor::clean_names() |>
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                           ordered = TRUE))
```


```{r}
bs |>
  filter(year == 2002) |>
  group_by(state) |>
  summarise(n_county = n_distinct(county)) |>
  filter(n_county >= 7)

bs |>
  filter(year == 2010) |>
  group_by(state) |>
  summarise(n_county = n_distinct(county)) |>
  filter(n_county >= 7)
```

In 2002, there are 6 states were observed at 7 or more location;
In 2010, there are 14 states were observed at 7 or more location.

```{r}
excellent = 
  bs |>
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarise(average = mean(data_value))
```


```{r}
excellent |>
  ggplot(aes(x = year, y = average, color = state)) +
  geom_point(alpha = .5) +
  geom_line(alpha = .5)
  labs(x = "Year", y = "Average Values for the Data",
       title = "Average values for Each State Across Years")
```


Based on the plot above, the Excellent rating across all states shows that there is a noticeable variance but hard to identify the overall trends.


```{r}
bs |>
  filter(year %in% c(2006,2010), state == "NY") |>
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_boxplot() + 
  facet_wrap(~year) +
  labs(x = "Responses",
       y = "Values",
       title = "Distribution of Data_value for Responses"
  )
```


Based on the plot above, 2006 and 2010 had relatively similar rating distributions, but 2010 overall contained larger range than 2006.


# Problem 3


```{r, load, tidy, and merge}
nc = 
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
  drop_na()

na = 
  read.csv("data/nhanes_accel.csv") |>
  janitor::clean_names()

nhanes =
  inner_join(na, nc, by = "seqn")
```

Load, tidy, and merge the datasets in to "nhanes" dataset.


```{r}
nhanes$seqn <- as.integer(nhanes$seqn)

nhanes$sex <- factor(nhanes$sex,
                     levels = c("1", "2"),
                     labels = c("male", "female"))

nhanes$education <- factor(nhanes$education,
                           levels = c("1", "2", "3"),
                           labels = c("less than high school", "high school equivalent", "more than high school"))

nhanes$age <- as.integer(nhanes$age)

nhanes$bmi <- as.numeric(nhanes$bmi)
```


```{r}
nhanes |>
  group_by(sex, education) |>
  summarise(count = n()) |>
  pivot_wider(names_from = education, values_from = count) |>
  knitr::kable()
```

The table above shows that there are 27 males and 28 females in the category less than high school,
there are 35 males and 23 females in the category high school equivalent, and there are 56 males 
and 59 females in the category more than high school.


```{r}
nhanes |>
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot(alpha = .5) +
  labs(x = "Education Level",
       y = "Age",
       fill = "Gender",
       title = "Age Distribution by Education and Gender"
  )
```


Based on the plot above, we can see that for category less than high school and more than high school, the overall age distribution for male and female are similar in these two groups, but for 
category high school equivalent male are younger than female, and for category less than high school has the smallest range for all these 3 categories.


```{r}
nhanes_total_activity <-
  nhanes |>
  mutate(total_activity = rowSums(across(min1:min1440), na.rm = TRUE))

nhanes_total_activity |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_smooth(method = lm, se = FALSE) +
  geom_point(alpha = .5) +
  facet_grid(~ education) +
  labs(x = "Age",
       y = "Tatal Activity",
       title = "Total Activity Over the Day by Education and Gender")
```


Based on the plot above, we can see that the overall association between total activity and age is negative, which means there is a decreased amount of total activity with age and the trend are the same for both gender. The total activity for categories high school equivalent and more than high school shows that female tend to have more total activity on average. And for category less than high school, totaly activity decrease the most as their age increase.


```{r}
nn <-
  nhanes |>
  pivot_longer(cols = starts_with("min"),
               names_to = "minute",
               values_to = "activity") |>
  mutate(minute = as.integer(gsub("min", "", minute)))
         
nn |>
  ggplot(aes(x = minute, y = activity, color = sex)) +
  geom_smooth(se = FALSE) +
  geom_line(alpha = .3) +
  facet_wrap(~ education, ncol = 1) +
  labs(x = "Minutes",
       y = "Activity",
       title = "24-hour Activity Time Courses for Each Education Level")
```


Based on the plot above, it is hard to tell the different of 24-hour activity time courses between male and female for all three education levels.
Overall, at about the 250th minute is the least amount of activity of the day.
Overall, there is higher amount of activity among female compare to male.